### 2018-04-09 latex/icpp18/introduction.tex
minor

NMF is widely used in data mining and machine learning as a dimension reduction and factor analysis method.
It is a natural fit for many real world problems as the non-negativity is inherent in many representations of real-world data and
the resulting low rank factors are expected to have a natural interpretation. The applications of NMF range from text mining \cite{pauca2004text},  computer vision \cite{hoyer2004non}, and bioinformatics \cite{kim2007sparse} to blind source separation  \cite{cichocki2009nonnegative}, unsupervised clustering \cite{kuang2012symmetric,kuang2013symnmf}  and many other areas.
In most real-world applications $m$ and $n$ are on the order of millions or more while $k$ is much smaller, on the order of tens to thousands.
Furthermore, data sets from these applications are often quite sparse and have highly irregular nonzero patterns.
NMF is an effective tool for analyzing such data in many applications, yet there is a lack of high performance algorithms and software that can compute it efficiently.
Our goal is in this paper is to fill in this gap by providing scalable parallel algorithms that employ efficient communication schemes and can scale to thousands of processes.

>>>>

NMF is widely used in data mining and machine learning as a dimension reduction and factor analysis method.
It is a natural fit for many real world problems as the non-negativity is inherent in many representations of real-world data and
the resulting low rank factors are expected to have a natural interpretation. The applications of NMF range from text mining \cite{pauca2004text},  computer vision \cite{hoyer2004non}, and bioinformatics \cite{kim2007sparse} to blind source separation  \cite{cichocki2009nonnegative}, unsupervised clustering \cite{kuang2012symmetric,kuang2013symnmf}  and many other areas.
In most real-world applications $m$ and $n$ are on the order of millions or more while $k$ is much smaller, on the order of tens to thousands.
Furthermore, data sets from these applications are often quite sparse and have highly irregular nonzero patterns.
NMF is an effective tool for analyzing such data in many applications, yet there is a lack of high performance algorithms and software that can compute it efficiently.
There are scalable tensor factorization algorithms resembling NMF kernels with sparse irregular computational patterns, yet they require an expensive partitioning step and do not necessarily yield the optimal performance in the context of NMF.
Our goal is in this paper is to fill in this gap by comparing the existing approaches and providing scalable parallel algorithms that employ efficient communication and partitioning schemes enabling scaling to thousands of processes.
